We are not the 99 percent: quantifying the atypical satellite distributions in the Local Group

### 2017-10-30 paper/paper.tex
mu

%=========================================================================
%		FRONT MATTER
%=========================================================================
\title[LG as an outlier]{We are not the 99 percent: the Local Group as an
  extreme outlier in ellipsoidal satellite distributions}
\author[J.E. Forero-Romero \& V. Arias]
{Jaime E. Forero-Romero $^{1}$ \thanks{je.forero@uniandes.edu.co},
Ver\'onica Arias$^1$\\
%%
$^1$ Departamento de F\'isica, Universidad de los Andes, Cra. 1
  No. 18A-10 Edificio Ip, CP 111711, Bogot\'a, Colombia \\
}

@@

\begin{abstract}
In this paper we quantify the joint spatial distribution of the
brightest satellites around the Milky Way (MW) and M31, the dominant
galaxies in the Local Group (LG).
We use the Illustris-1 and ELVIS simulations to quantify the
expected deviations from statistically spherical satellite
distributions in the Lambda Cold Dark Matter paradigm. 
We estimate that only $0.05\%$ to $0.25\%$ of the pairs with isolation
and dynamical characteristics similar to the LG are expected to have
satellite distributions with the same degree of deviation from
sphericity.
This makes the LG a $3\sigma$ outlier, with all the weight lying on
the extremly planar (width of $22\pm 2$kpc and mayor-to-minor axis
ration of $c/a=0.53\pm0.03$) satellite distribution in the MW. 
\end{abstract}

@@

Figure \ref{fig:scatter_width} summarizes the main results for the plane
width distributions.
The panel on the left compares the results for the MW and M31
observations against its randomized version. 
The most interesting outcome is that the MW plane width is smaller
than $95\%$ of the planes computed from the randomized distribution,
while the M31 plane width is consistent with the same distribution.

@@

M31 observations

>>>>

%=========================================================================
%		FRONT MATTER
%=========================================================================
\title[LG as an outlier]{We are not the 99 percent: quantifying the
  atypical satellite distributions in the Local Group}
\author[J.E. Forero-Romero \& V. Arias]
{Jaime E. Forero-Romero $^{1}$ \thanks{je.forero@uniandes.edu.co},
Ver\'onica Arias$^1$\\
%%
$^1$ Departamento de F\'isica, Universidad de los Andes, Cra. 1
  No. 18A-10 Edificio Ip, CP 111711, Bogot\'a, Colombia \\
}

@@

\begin{abstract}
In this paper we quantify the joint spatial distribution of the
brightest satellites around the Milky Way (MW) and M31, the dominant
galaxies in the Local Group (LG).
We use the Illustris-1 and ELVIS simulations to quantify the
expected deviations from statistically spherical satellite
distributions in the Lambda Cold Dark Matter paradigm. 
We estimate that only $0.05\%$ to $0.25\%$ of the pairs with isolation
and dynamical characteristics similar to the LG are expected to have
satellite distributions with the same degree of deviation from
sphericity.
This makes the LG a $3\sigma$ outlier, with all the weight lying on
the extremly planar MW satellite distribution.
\end{abstract}

@@

Figure \ref{fig:scatter_width} summarizes the results for the plane
width distributions.
The panel on the left compares the results for the MW and M31
observations against its randomized version. 
The most interesting outcome is that the MW plane width is smaller
than $95\%$ of the planes computed from the randomized distribution,
while the M31 plane width is consistent with the same distribution.

@@

Figure \ref{fig:scatter_mu} summarizes the Illustris results for the
alignments between the vector normal to the plane and the vector
connecting the two dominant galaxies.
The corresponding results for the ELVIS data are shown in Figure
\ref{fig:scatter_elvis}.

The first striking impression from the left panel in Figure
\ref{fig:scatter_mu} is that the observations are cornered in
a region of strong alignment, $\mu\approx 1$, for both galaxies. 
In other words, the planes are almost perpendicular to the direction
connecting the two galaxies. 
The result for the spherical randomization corresponds to a uniform
distribution $0\leq \mu\leq 1$ as shown in the same panel.

The middle panel tells a similar story for the results from
Illustris. That is, there isn't a prefered alignment direction from
the simulation. This also holds for ELVIS data.
We quantify this visual impresion by performing a K-S test where the
null hypothesis corresponds to an uniform distribution. 
The results from the K-S test do not allow us to discard the null
hypotesis.

This information allows us to properly understand the results on the
right panel in Figure \ref{fig:scatter_mu}.
That panel shows the results for the normalized values to the mean and
standard deviation computed from spherically randomized
ditributions.
In this case the LG $\mu$ value has the same probability as any other value
(because it is consistent with an uniform distribution) and the apparent
large distance from the Illustris data must not be interpreted as an
the LG having an atypical alignment.

We defer to the Section \ref{sec:discussion} our discussion on how
this
